---
title: "Workaround for long run"
output: html_document
date: "2026-03-17"
---


Sometime when analysing large phenotype matrices (say many ATAC seq conditions) in a cluster
the multFSuSiE fitting procedure main take longer than the allowed wall time.

Here we showcase how to use a given fit that has not reached convergence and to 
continue iterating. This allow you to save the fit every say 20 iterations and 
to restart a job at the exact same step as when the data point was fitted

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Here we simulate a relatively simple example in which SNP can affect 2 functional/spatially structure traits and 3 univariate traits.
 
```{r data_generation}
library(mvf.susie.alpha)
library(susieR)
set.seed(1)

N <- 100 # Sample size 
L <- 2 # Number of effects
list_lev_res <- list(7, 5)
# Two functional phenotypes, one of length 2^5 and one of length 2^6
n_univ <- 3 # 3 univariate phenotypes
eff <- list()
for (l in 1:L) { # Simulate the multi-trait effect
  eff[[l]] <- simu_effect_multfsusie(list_lev_res=list_lev_res,
                                     n_univ=n_univ, output_level=2)
}

Y_f1 <- matrix(rnorm((2^list_lev_res[[1]]) * N, sd=.51), nrow=N)
Y_f2 <- matrix(rnorm((2^list_lev_res[[2]]) * N, sd=.51), nrow=N)
Y_u <- matrix(rnorm(n_univ * N, sd=1), nrow=N)
# Genotype matrix
G <- N3finemapping$X[1:100, ]

true_pos <- sample(1:ncol(G), L) # Actually causal columns/SNPs

for (i in 1:N) {
  for (l in 1:L) {
    Y_f1[i,] <- Y_f1[i,] + eff[[l]]$func_effect[[1]]$sim_func * G[i, true_pos[[l]]]
    Y_f2[i,] <- Y_f2[i,] + eff[[l]]$func_effect[[2]]$sim_func * G[i, true_pos[[l]]]
    Y_u[i,]  <- Y_u[i,]  + eff[[l]]$univ_effect * G[i, true_pos[[l]]]
  }
}

Y_f <- list(Y_f1, Y_f2)
Y <- list(Y_f=Y_f, Y_u=Y_u) # Preparing data
pos <- list(pos1=1:ncol(Y$Y_f[[1]]), pos2=1:ncol(Y$Y_f[[2]]))

```


We showcase that storing the multfsusie object and them re-injecting it in the fit
lead to the same results as the full procedure.

Note the that parameter maxit controls the total number of iterations (default 100) throughout your iterations
whereas the parameter max_step controls the total number of iterations  (default 100) in a given call of the multfsusie function.

## Including Plots

You can also embed plots, for example:

```{r example}


# If your signal is sampled between 1 and 64

m0 <- multfsusie(Y=Y, X=G, pos=pos, L=3, max_step = 4)
m0$check


```

So you get a wrning message to tell you that the fitting stopped early and the part m0$check gives
you the difference in fitted alpha values between iteration.

Let's now iterate the fit

```{r}
m1=m0
for ( k in 1:5){
  if (m1$check>1e-6) { #default convergence criterion 
     m1 <- multfsusie(Y=Y, X=G, pos=pos, L=3, max_step = 4,
                   multfsusie.obj = m1)
  }
 
}

```

Well done now lets fit everything in one go for the sake of comparison
```{r}

mfull <- multfsusie(Y=Y, X=G, pos=pos, L=3 )
```

```{r}
mfull$check
m1$check
```




 
```{r}
par(mfrow=c(2,2))
plot( mfull$fitted_func[[1]][[1]], ylab="effect estimate")
lines(m1$fitted_func[[1]][[1]])
plot( mfull$fitted_func[[1]][[1]],m1$fitted_func[[1]][[1]], 
      ylab="effect estimate iterated", xlab="effect estimate full")
abline(a=0,b=1)
plot( mfull$fitted_func[[2]][[1]], ylab="effect estimate")
lines(m1$fitted_func[[2]][[1]])
plot( mfull$fitted_func[[2]][[1]],
      m1$fitted_func[[2]][[1]], 
      ylab="effect estimate iterated", xlab="effect estimate full")
abline(a=0,b=1)
```

